---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---

##Library
```{r}
rm(list=ls())
library(tidyverse)
library(readr)
library(lubridate)
library(dplyr)
library(lme4)
library(tidyverse)
```

##Data
```{r}
BDA <- read_csv("Bressoux.csv")
```


##code
```{r}
BDA <-
BDA %>%
  rename(Studentid = NUMELEVE,
         Classid = CLASSE2,
         MathScore = MATH4,
         LabourFather = ouvr,
         NStudentInClass = NBEL2)
```
Variable of Interest:

  Studentid (renamed from NUMELEVE) - Student id (no numerical intepretation, value indicates index of different student)
  
  Classid (renamed from CLASSE2) - Class id (no numerical intepretation, value indicates index of different class)
  
  MathScore (renamed from MATH4) - End-of-year score at standardized maths test (numerically 
meaningful, for example, the larger the value, the higher the score)
  
  LabourFather (renamed from ouvr) - whether the student's father is laborer or not (1 = yes, 0 = no, with no numerical intepretation)
  
  NStudentInClass (renamed from NBEL2) - Number of students in the class (numerically 
meaningful, for example, the larger the value, the greater the class population)
  
  
Research Questions: 
  
   What is the association between whether the student's father is laborer or not (1 = yes, 0 = no, for variable LabourFather) and End-of-year score at standardized maths test (MathScore Variable)
   
   Does the association differ based on the number of students in the class (NStudentInClass Variable)
   

Level 1 observation unit - Student, differentiated by Studentid 

Level 2 observation unit - class, differentiated by Classid 

Response variable - MathScore (Level 1 variable but not explanatory variable)

Level 1 explanatory variables -  LabourFather

Level 2 explanatory variables -  NStudentInClass

Fixed effects:

  - LabourFather (1 = yes, 0 = no)
  
  - NStudentInClass 
  
  
Random effects: 

  - Classid (Class id)
  
```{r}
BDA_MATH <- ggplot(data=BDA, mapping=aes(x=MathScore))+
  geom_bar(binwidth = 2, color="black",fill="white")+
  ylim(0,30)+
  labs(x="Students Math Score",
       y="Number of Students")
BDA_MATH
```
This graph is bell shaped around value of 0.

```{r}
BDA_L <- ggplot(data=BDA, mapping=aes(x=LabourFather))+
  geom_bar(color="black",fill="white")+
  labs(x="Labour Father or not",
       y="Frequency")
  
BDA_L
table(BDA$LabourFather)
```
In the data, We have 415 students who doesn't have a labour father and the rest 194 students do have a labour father. 1 is students who have a labour father and 0 is students who does not have a labour father.

```{r}
StudentM <- ggplot(data = BDA, mapping = aes(y = Studentid, x = MathScore)) + 
  geom_boxplot() +
  labs(x = "MathScore",
       y = "Students number") + 
  facet_wrap(  ~ LabourFather)
print(StudentM + ggtitle("Mathscore for students with or withour a labour father, 0=NO,1=YES"))

ClassM <- ggplot(data = BDA, mapping = aes(y = Classid, x = MathScore)) + 
  geom_boxplot() +
  labs(x = "MathScore",
       y = "Class number") + 
  facet_wrap(  ~ LabourFather)
print(ClassM + ggtitle("Mathscore for class with or withour a labour father, 0=NO,1=YES"))
```
There is not much difference on this box plot comparing students father is labour or not. The Math Score will not change significantly based on students who have a labour father or not. 

```{r}
BDA %>% 
  group_by(Studentid) %>%
  summarize(meanmath = mean(MathScore, na.rm = TRUE),
          sdmath = sd(MathScore, na.rm = TRUE))

BDA %>% 
  group_by(Classid) %>%
  summarize(meanmath = mean(MathScore, na.rm = TRUE),
          sdmath = sd(MathScore, na.rm = TRUE))
```
Here we summarize mean and sd MathScore by studentid or classid. 

```{r}
summary(BDA$MathScore)
summary(BDA$NBEL2)
```
Here we generate the five number summary.

```{r}
ggplot(data = BDA_L, mapping = aes(y = MathScore, x = LabourFather)) + 
  geom_boxplot() +
  labs(x = "LabourFather or not",
       y = "MathScore") + 
  facet_wrap(  ~ Classid)
```

```{r}
ggplot(data = BDA, mapping = aes(x = NBEL2)) + 
  geom_histogram(color = "black", fill = "white", binwidth = 2) +
  labs(x = "Number of students in class")
```
Based on this graph, we know most student have math score between valye of -1 and 1.

## Model Selection

$$
\text{Random Intercepts model} \\
Y_{ij} = a_i + \epsilon_{ij} \\  
a_i = \alpha_0 + \mu_i \\
\text{where } \epsilon_{ij} \text{~N(0,} \sigma_u^2 )
$$

```{r}
model1 <- lmer(MathScore ~ 1 + (1 | Classid), data = BDA, REML = TRUE)
summary(model1)
coef(model1)
```

$$
\text{Random Intercepts and slopes model based on inclusion of LabourFather Variable} \\
Y_{ij} = a_i + b_i LabourFather_i + \epsilon_{ij} \\
a_i = \alpha_0 + \mu_i \\
b_i = \beta_0 + v_i \\
\text{where } \epsilon_{ij} \text{~N(0,} \sigma_u^2 )
$$


```{r}
model2 <- lmer(MathScore ~ LabourFather + (1 + LabourFather | Classid), data = BDA, REML = TRUE)
summary(model2)
coef((model2))
```

$$
\text{Random Intercepts and slopes model based on inclusion of LabourFather and NStudentInClass Variables} \\
Y_{ij} = a_i + b_i LabourFather_i + \epsilon_{ij} \\
a_i = \alpha_0 + \alpha_1 NStudentInClass_i +  \mu_i \\
b_i = \beta_0 + \beta_1 NStudentInClass_i + v_i \\
\text{where } \epsilon_{ij} \text{~N(0,} \sigma_u^2 )
$$


```{r}
model3 <- lmer(MathScore ~ NStudentInClass + LabourFather + NStudentInClass:LabourFather + (1 + LabourFather | Classid), data = BDA, REML = TRUE)
summary(model3)
coef(model3)
```

```{r}
model1t <- lmer(MathScore ~ 1 + (1 | Classid), data = BDA, REML = FALSE)
model3t <- lmer(MathScore ~ NStudentInClass + LabourFather + NStudentInClass:LabourFather + (1 + LabourFather | Classid), data = BDA, REML = FALSE)
anova(model1t, model3t, test = "chisq")
```

```{r}
BDA_M4 <- ggplot(data=BDA, mapping=aes(x=MATH4))+
  geom_bar(binwidth = 2, color="black",fill="white")+
  ylim(0,30)+
  labs(x="Negative Affect",
       y="Number of Musiciams")
BDA_M4
```


